---
output: 
  github_document:
    toc: TRUE
    toc_depth: 3
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>", 
  message = F,
  warning = F,
  error = T,
  cache = F
)
```

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

# {ggregions}

R Medicine Talk [slides](https://evamaerey.github.io/mytidytuesday/2026-04-03-r-medicine-ggregions/r-medicine-ggregions.html#1)

COWY-ASA Talk 2026-05-08 [slides](https://evamaerey.github.io/mytidytuesday/2026-04-03-r-medicine-ggregions/asa-cowy-ggregions.html#1)

Note: foundation work for this package has been conducted in many experiments 
([1](https://evamaerey.github.io/mytidytuesday/2025-09-15-make_constructor_sf/make_constructor_sf.html),
[2](https://evamaerey.github.io/mytidytuesday/2024-07-12-sf-experiment-fastStat/sf-experiment-fastStat.html),
[3](https://evamaerey.github.io/mytidytuesday/2024-06-13-usmaps-ggusa/usmaps-ggusa.html ),
[4](https://evamaerey.github.io/mytidytuesday/2024-06-13-usmaps-ggusa/usmaps-ggusa-Stat-and-geom-sf.html ),
[5](https://evamaerey.github.io/mytidytuesday/2024-04-08-statsf_errors/statsf_errors.html ),
[6](https://evamaerey.github.io/mytidytuesday/2023-03-12-ggbrain-seg-sf/ggbrain_seg_sf.html ),
[7](https://evamaerey.github.io/mytidytuesday/2023-03-12-ggbrain-seg-sf/ggbrain_seg_dk_sf.html ),
[8](https://evamaerey.github.io/mytidytuesday/2023-03-10-ggfips/ggfips_w_sf.html ),
[9](https://evamaerey.github.io/mytidytuesday/2023-03-10-ggfips/ggfips.html ),
[10](https://evamaerey.github.io/mytidytuesday/2023-03-10-brain2/brain2.html ),
[11](https://evamaerey.github.io/mytidytuesday/2023-03-09-nc-fips/nc-fips.html ),
[12](https://evamaerey.github.io/mytidytuesday/2023-03-07-extending-your-ability/extending_your_ability.html ),
[13](https://evamaerey.github.io/mytidytuesday/2023-03-06-us-states/us_states.html ) ),
 [a number of discussions](https://github.com/search?q=repo%3Aggplot2-extenders%2Fggplot-extension-club+sf2stat&type=discussions) and in conversation based on my [talk](https://evamaerey.github.io/mytidytuesday/2023-04-13-ggcircle-pack-talk/circle_pack_and_beyond_talk.html#86) in the ggplot2 extenders group,
and a couple of experimental packages ([sf2stat](https://github.com/EvaMaeRey/sf2stat) and [ggsomewhere](https://github.com/EvaMaeRey/ggsomewhere)).

For an example of how the ggregions utility might be used in a specific geographic context, see [ggusmap](https://github.com/EvaMaeRey/ggusmap), which uses the superb data prepared by [Paolo Di Lorenzo](https://github.com/pdil) for the usmap and usmapdata packages.  



# Motivations

> I am told there are people who do not care for maps, and I find it hard to believe. - Robert Louis Stevenson

> 'My teacher, Mr. Jayson, says a map is a picture of someplace from above.  It's like flying over that spot in an airplane.' - 'Lisa' in Loreen Leedy's in 'Mappying Penny's World'

> I hate when I work with spatial data that so much of my mindshare goes to thinking about 'how do I wrangle this this map and what on earth is going on' versus the actual meaning behind what I'm trying to map. - Emily Riederer

Maps are some of the most compelling and intuitive data visualizations, and they are some of the first visualizations people will come across in their lives.  There is great support for mapping in R and in the data visualization library {ggplot2} which supports many R packages.  

However, producing a simple map, like a choropleth, can feel much harder than producing other plots, like a scatterplot for example in ggplot2.  A scatter plot is produced by:

```r 
ggplot(scatter_data) + # data
   aes(x = x, y = y) + # position and other visual channel mapping
   geom_point()        # layer 
```

In other words, you might say, 'used this data, and points, and the positions are defined by x and y'

This proposal posits that by investing in methods to easily create new geo-specific ggplot2 layer extensions (geom_*s), producing maps with ggplot2 can become as intuitive as creating other plots. About specifying a map, you might say, 'use this data, draw regions, position and color fill according to var1 and var2.'

The proposed regions mapping might look something like this:

```r
ggplot(mapping_data) +              # data
   aes(region = var1_region_id,  # position mapping 
       fill = var2) +               # and other visual channel mapping
   geom_region()                    # layer
```

> Your API feels like it actually puts the "question" at the center.. - Emily Riederer



## Status Quo (without ggregions): Leave the data manipulation to the user, or wrapper APIs

```{r, message=F, warning=F, fig.show='hide'}
library(tidyverse)
library(ozmaps)
sf_oz <- ozmap("states")
```

```{r}
# some data that we want to viz
au_states <- tribble(~state, ~pop,
        "Victoria", 1,
        "Queensland", 2,
        "New South Wales", 3,
        "Western Australia", 4,
        "Northern Territory", 5,
        "Tasmania", 6,
        "South Australia", 7,
        "Australian Capital Territory", 8,
        "Other Territories", 9
        )

sf_oz |>
  full_join(au_states, by = join_by(NAME == state )) |> 
  ggplot() + 
  aes(geometry = geometry, 
      fill = pop) + 
  geom_sf()
```

# What {ggregions} delivers



## Interface #1. Standard interface for visualizing regions and locales, `aes(region = my_var) + geom_region()` where scope of interest is declared via `options()`.

```{r}
library(tidyverse)
```

<details>

```{r}
# this should probably be changed...
ref_data_us <- usmapdata::us_map() |>
  dplyr::select( 
    state_abbr = abbr, # first variable will also be 'id' var
    state_name = full, 
    state_fips = fips,
    geometry = geom,  # one column 'geometry' is required
         )  

usethis::use_data(ref_data_us, overwrite = T)
```


```{r geom_region}
st_crs_mod <- function(ref_data){

  crs <- sf::st_crs(ref_data)
  
  if(is.na(crs)){NULL}else{crs}

}


compute_panel_regions <- function (data, scales, ref_data, keep = NULL, drop = NULL, stamp = F) {
  
    ref_data$id <- ref_data[1][[1]]
    
    if (!is.null(keep)) {
        ref_data <- dplyr::filter(ref_data, id %in% keep)
    }
    
    if (!is.null(drop)) {
        ref_data <- dplyr::filter(ref_data, !(id %in% drop))
    }
    
    if (!stamp) {
    
          ref_data_long <- ref_data |> 
            mutate(across(-geometry, as.character)) |>
            pivot_longer(cols = -c(geometry, id), 
                         names_to = ".id_type", 
                         values_to = "region")

          # check unique in the future
          length(ref_data_long$region) == length(unique(ref_data_long$region))
    
          ref_data_long <- ref_data_long |> 
            ggplot2::StatSf$compute_panel(coord = ggplot2::CoordSf) |>
            ggplot2::StatSfCoordinates$compute_group(coord = ggplot2::CoordSf)
        
          out <- dplyr::inner_join(ref_data_long, data, by = join_by(region)) 
          
          ref_data |> sf::st_drop_geometry() |>
            inner_join(out, by = join_by(id))
          
    }
    
    else {
      
        ref_data |> 
            ggplot2::StatSf$compute_panel(coord = ggplot2::CoordSf) |>
            ggplot2::StatSfCoordinates$compute_group(coord = ggplot2::CoordSf)
      
    }
    
}


StatRegion <- ggplot2::ggproto("StatRegion", ggplot2::Stat,
                      compute_panel = compute_panel_regions,
                      required_aes = "region",
                      default_aes = ggplot2::aes(label = ggplot2::after_stat(id),
                                                 geometry = geometry))

StatRegionStamp <- ggplot2::ggproto("StatRegionStamp", ggplot2::Stat,
                      compute_panel = compute_panel_regions,
                      default_aes = ggplot2::aes(label = ggplot2::after_stat(id),
                                                 geometry = geometry))

#' @export
geom_region <- function (mapping = aes(), data = NULL, 
                         stat = StatRegion,
                         position = "identity", 
                         na.rm = FALSE, show.legend = NA, 
                         inherit.aes = TRUE, ref_data = getOption("ggregions.ref.regions", ref_data_us), ...){
    c(layer_sf(geom = GeomSf, data = data, mapping = mapping, 
        stat = stat, position = position, show.legend = show.legend, 
        inherit.aes = inherit.aes, params = rlang::list2(na.rm = na.rm, ref_data = ref_data,
            ...)), coord_sf(crs = st_crs_mod(ref_data))
      )
}


#' @export
stamp_region <- function (mapping = aes(), stat = StatRegionStamp, 
                          position = "identity", 
                          na.rm = FALSE, show.legend = NA, 
                          inherit.aes = FALSE, ref_data = getOption("ggregions.ref.regions", ref_data_us), ...) 
{
    c(layer_sf(geom = GeomSf, data = ref_data, mapping = mapping, 
        stat = stat, position = position, show.legend = show.legend, 
        inherit.aes = inherit.aes, params = rlang::list2(na.rm = na.rm, 
                                                         ref_data = ref_data, stamp = T,
            ...)), coord_sf(crs = st_crs_mod(ref_data))
      )
}


GeomSfBorder <- ggplot2::ggproto("GeomSfBorder", ggplot2::GeomSf, 
                        default_aes = ggplot2::GeomSf$default_aes |>
                          modifyList(
                            ggplot2::aes(fill = "transparent", 
                                         linewidth = ggplot2::from_theme(linewidth*1.5)),
                            keep.null = T)) 

#' @export
geom_region_border <- function (mapping = aes(), data = NULL, stat = StatRegion, position = "identity", 
    na.rm = FALSE, show.legend = NA, inherit.aes = TRUE, ref_data = getOption("ggregions.ref.regions", ref_data_us), ...) 
{
    c(layer_sf(geom =  GeomSfBorder, data = data, mapping = mapping, 
        stat = stat, position = position, show.legend = show.legend, 
        inherit.aes = inherit.aes, params = rlang::list2(na.rm = na.rm, ref_data = ref_data, fill = "transparent",
            ...)), coord_sf(crs = st_crs_mod(ref_data))
      )
}


#' @export
stamp_region_border <- function (mapping = aes(), data = ref_data, 
                          stat = StatRegionStamp, position = "identity", 
    na.rm = FALSE, show.legend = NA, inherit.aes = FALSE, ref_data = getOption("ggregions.ref.regions", ref_data_us), stamp = TRUE, ...) 
{
    c(layer_sf(geom = GeomSfBorder, data = data, mapping = mapping, 
        stat = stat, position = position, show.legend = show.legend, 
        inherit.aes = inherit.aes, params = rlang::list2(na.rm = na.rm, ref_data = ref_data, stamp = stamp, fill = "transparent",
            ...)), coord_sf(crs = st_crs_mod(ref_data))
      )
}

#' @export
geom_region_text <- function (mapping = aes(), data = NULL, stat = StatRegion, position = "identity", 
    na.rm = FALSE, show.legend = NA, inherit.aes = TRUE, ref_data = getOption("ggregions.ref.regions", ref_data_us), ...) 
{
    c(layer_sf(geom = GeomText, data = data, mapping = mapping, 
        stat = stat, position = position, show.legend = show.legend, 
        inherit.aes = inherit.aes, params = rlang::list2(na.rm = na.rm, ref_data = ref_data,
            ...)), coord_sf(crs = st_crs_mod(ref_data))
      )
}


#' @export
stamp_region_text <- function (mapping = aes(), data = ref_data, 
                          stat = StatRegionStamp, position = "identity", 
    na.rm = FALSE, show.legend = NA, inherit.aes = FALSE, ref_data = getOption("ggregions.ref.regions", ref_data_us), stamp = TRUE, ...) 
{
    c(layer_sf(geom = GeomText, data = data, mapping = mapping, 
        stat = stat, position = position, show.legend = show.legend, 
        inherit.aes = inherit.aes, params = rlang::list2(na.rm = na.rm, ref_data = ref_data, stamp = stamp,
            ...)), coord_sf(crs = st_crs_mod(ref_data))
      )
}

#' @export
geom_region_label <- function (mapping = aes(), data = NULL, stat = StatRegion, position = "identity", 
    na.rm = FALSE, show.legend = NA, inherit.aes = TRUE, ref_data = getOption("ggregions.ref.regions", ref_data_us), ...) 
{
    c(layer_sf(geom = GeomLabel, data = data, mapping = mapping, 
        stat = stat, position = position, show.legend = show.legend, 
        inherit.aes = inherit.aes, params = rlang::list2(na.rm = na.rm, ref_data = ref_data,
            ...)), coord_sf(crs = st_crs_mod(ref_data))
      )
}

#' @export
stamp_region_label <- function (mapping = aes(), data = ref_data, 
                          stat = StatRegionStamp, position = "identity", 
    na.rm = FALSE, show.legend = NA, inherit.aes = FALSE, ref_data = getOption("ggregions.ref.regions", ref_data_us), stamp = TRUE, ...) 
{
    c(layer_sf(geom = GeomLabel, data = data, mapping = mapping, 
        stat = stat, position = position, show.legend = show.legend, 
        inherit.aes = inherit.aes, params = rlang::list2(na.rm = na.rm, ref_data = ref_data, stamp = stamp,
            ...)), coord_sf(crs = st_crs_mod(ref_data))
      )
}

```


</details>

### North Carolina example

First, the authors will prepare a reference dataset that contains a [geometry](https://ggplot2.tidyverse.org/reference/ggsf.html#geometry-aesthetic), sfc-containing column as well as columns of allowed identifiers for the regions of interest.  For example, the identifiers from in `nc_ref` are `county_name` and `county_fips`. 

```{r, fig.show='hold'}
# Interface #1

# Step 1.a Create reference data set
nc <- sf::st_read(system.file("shape/nc.shp", package="sf"))

nc_ref <- nc |> 
  select(county_name = NAME, 
         county_fips = FIPS, 
         geometry)
```


```{r}
# Step 1.b Declare reference data in options
options(ggregions.ref.regions = nc_ref)
```


```{r}
# Step 2. Plot! Use aes(region = my_var_regional_id) + geom_region()
tribble(~county, ~ind_going,
        "Ashe",           1,
        "Northampton",    0
        ) |>
  ggplot() + 
  aes(region = county,
      fill = ind_going) +
  stamp_region()

last_plot() +
  geom_region()


```


### Australia example

```{r, fig.show='hide'}
## Step 1a Prep reference data
library(ozmaps)
sf_oz <- ozmap("states")

australia_state_ref <- sf_oz |>
  select(state_name = NAME, geometry)
```

```{r}
## Step 1b declare reference data
options(ggregions.ref.regions = australia_state_ref)
```

```{r}
## Step 2 plot with geom_region and frinew

tribble(~state, ~pop,
        "Victoria", 1,
        "Queensland", 2,
        "New South Wales", 3,
        "Western Australia", 4,
        "Northern Territory", 5,
        "Tasmania", 6,
        "South Australia", 7,
        "Australian Capital Territory", 8,
        "Other Territories", 9
        ) |> 
  ggplot() + 
  aes(region = state, fill = pop) + 
  geom_region() +
  geom_region_border(keep = "Tasmania", color = "cadetblue2")

library(ggplot2)
ggplot() + 
  stamp_region() + 
  stamp_region(keep = "Victoria", fill = "blue") + 
  stamp_region_border(keep = "Western Australia", 
                 color = "orange") + 
  stamp_region_text(keep = "Queensland")
```



## Interface #2.  use write_*() functions to specify reference data for layers.

<details>

```{r write_geom_region, eval = T}
#' @export
write_geom_region <- function(ref_data){

  modified_fun <- geom_region

  formals(modified_fun)$ref_data <- substitute(ref_data)

  return(modified_fun)

}



#' @export
write_stamp_region <- function(ref_data){

  modified_function <- stamp_region

formals(modified_function)$ref_data <- substitute(ref_data)
# formals(modified_function)$data <- substitute(ref_data)


return(modified_function)

}


#' @export
write_geom_region_border <- function(ref_data){

  modified_fun <- geom_region_border

  formals(modified_fun)$ref_data <- substitute(ref_data)

  return(modified_fun)

}

#' @export
write_stamp_region_border <- function(ref_data){

  modified_fun <- geom_stamp_border

  formals(modified_fun)$ref_data <- substitute(ref_data)

  return(modified_fun)

}


#' @export
write_geom_region_text <- function(ref_data){

  modified_function <- geom_region_text

  formals(modified_function)$ref_data <- substitute(ref_data)

return(modified_function)

}


#' @export
write_stamp_region_text <- function(ref_data){

  modified_function <- stamp_region_text

  formals(modified_function)$ref_data <- substitute(ref_data)

  return(modified_function)

}


#' @export
write_geom_region_label <- function(ref_data){

  modified_function <- geom_region_label

  formals(modified_function)$ref_data <- substitute(ref_data)

return(modified_function)

}


#' @export
write_stamp_region_label <- function(ref_data){

  modified_function <- stamp_region_label

  formals(modified_function)$ref_data <- substitute(ref_data)

  return(modified_function)

}
```

</details>


### North Carolina example



```{r}
# Step 1.a prep ref data
nc_ref <- nc |> 
  select(county_name = NAME, 
         county_fips = FIPS, 
         geometry)

# Step 1.b
geom_county <- ggregions::write_geom_region(ref_data = nc_ref)
stamp_county <- ggregions::write_stamp_region(ref_data = nc_ref)


tribble(~county, ~ind_going,
        "Ashe",           1,
        "Northampton",    0
        ) |>
  ggplot() + 
  aes(region = county,
      fill = ind_going) + 
  stamp_county() + 
  geom_county()


tribble(~county, ~ind_going,
        "Ashe",           1,
        "Northampton",    0
        ) |> 
  rename(region = county) |>
  compute_panel_regions(ref_data = nc_ref)

```


### US example


```{r, eval = T}
us_income <- tidyr::us_rent_income |> 
  filter(variable == "income") |>
  rename(income = estimate)

head(us_income)
```

```{r, eval = T}
# step 1.a prepare ref data 

us_states_ref <- usmapdata::us_map() |>
  select( 
    state_abbr = abbr, # first variable will also be 'id' var
    state_name = full, 
    state_fips = fips,
    geometry = geom,  # one column 'geometry' is required
         )  

# step 1b write locked-in ref_data
geom_state        <- write_geom_region(ref_data = us_states_ref)
geom_state_text   <- write_geom_region_text(ref_data = us_states_ref)
geom_state_label  <- write_geom_region_label(ref_data = us_states_ref)
geom_state_border <- write_geom_region_border(ref_data = us_states_ref)

stamp_state        <- write_stamp_region(ref_data = us_states_ref)
stamp_state_text   <- write_stamp_region_text(ref_data = us_states_ref)
stamp_state_label  <- write_stamp_region_label(ref_data = us_states_ref)
stamp_state_border <- write_stamp_region_border(ref_data = us_states_ref)

# step 2 plot
head(us_income)

ggplot(data = us_income) + 
  aes(region = NAME, 
      fill = income) + 
  geom_state() + 
  geom_state_border(keep = "North Carolina")

```






# Minimal Packaging




### Do once

```{r, eval = F}
devtools::create(".")
usethis::use_lifecycle_badge("experimental")
```

### Do manually and in readme

```{r}
knitrExtra::chunk_to_dir("geom_region")
knitrExtra::chunk_to_dir("StatRegion0")
knitrExtra::chunk_to_dir("geom_region0")
knitrExtra::chunk_to_dir("write_geom_region")

usethis::use_package("ggplot2")
usethis::use_package("sf")
usethis::use_package("dplyr")

# rm(list = ls())
devtools::document()
```


### Do manually (time consuming)

```{r, eval = F}
devtools::check(".")
devtools::install(pkg = ".", upgrade = "never")
```




# Previous, recent implementation and experiments, see [source](https://github.com/EvaMaeRey/ggregions/blob/main/README.Rmd)

```{r}
knitr::knit_exit()
```



<details>








## Alternative: Toward region-specific layer-writing 


### Step 1. Define compute

We'll look at the Australian states case as we go, but the ref_data argument will keep thing generic.

```{r}
australia_state_ref <- sf_oz |>
  select(state_name = NAME)
```


```{r, eval = F}
compute_panel_regions <- function(data, scales, ref_data, keep = NULL, drop = NULL, stamp = F){

  ref_data$id <- ref_data[1][[1]]

  if(!is.null(keep)){ref_data <- ref_data |> dplyr::filter(id %in% keep)}
  if(!is.null(drop)){ref_data <- ref_data |> dplyr::filter(!(id %in% drop))}

  ref_data <- ref_data |> 
     ggplot2::StatSf$compute_panel(coord = ggplot2::CoordSf) |> # add bounding boxes xmin xmax etc
     ggplot2::StatSfCoordinates$compute_group(coord = ggplot2::CoordSf) # add x and y centers

  if(!stamp){ ref_data |> dplyr::inner_join(data) } else {ref_data}

}
```


### Test Compute

```{r}
au_states |> 
  rename(state_name = state) |>
  compute_panel_regions(ref_data = australia_state_ref)
```

To do/consider

- [ ] warnings about not plot data that fails to find a match in ref data
- [ ] allowing multiple geo reference data sets with different resolutions?
- [ ] more flexibility on specifying the id column (now it is the first column in the dataset)
- [ ] other?

### Step 2. Define Stat

```{r StatRegion0}
# compute just reproduced verbatim from above for packaging purposes
#' Compute panel regions
#'
#' @description
#' A short description...
#'
#' @param data A data frame.
#' @param scales Scales.
#' @param ref_data A reference data frame.
#' @param keep A character vector of IDs to keep. Optional.
#' @param drop A character vector of IDs to drop. Optional.
#' @param stamp A single logical value. Optional.
#'
#' @returns
#' A data frame. If `stamp` is `FALSE`, the processed `ref_data` inner-joined with `data`.
#' Otherwise, the processed `ref_data`.
#'
#' @export
compute_panel_regions0 <- function(data, scales, ref_data, keep = NULL, 
                                  drop = NULL, stamp = F){

  ref_data$id <- ref_data[1][[1]]

  if(!is.null(keep)){ref_data <- ref_data |> dplyr::filter(id %in% keep)}
  if(!is.null(drop)){ref_data <- ref_data |> dplyr::filter(!(id %in% drop))}

  ref_data <- ref_data |> 
    ggplot2::StatSf$compute_panel(coord = ggplot2::CoordSf) |>
    ggplot2::StatSfCoordinates$compute_group(coord = ggplot2::CoordSf)

  if(!stamp){ ref_data |> dplyr::inner_join(data) } else { ref_data }

}


StatRegion0 <- ggplot2::ggproto("StatRegion0",
                      ggplot2::Stat,
                      compute_panel = compute_panel_regions0,
                      default_aes = 
                        ggplot2::aes(label = ggplot2::after_stat(id)))
```

### To do/consider

- [ ] Stat won't warn about required_aes, since none are declared.  We anticipate allowed aes to be quite varied.  How otherwise can this be addressed?
- [ ] Think through, experiment if compute_group might also work, and if it is preferable

### Test Stat

```{r}
ggplot(au_states) +
  aes(state_name = state) +
  geom_sf(stat = StatRegion0, ref_data = australia_state_ref) + 
  geom_text(stat = StatRegion0, ref_data = australia_state_ref) +
  aes(fill = pop)
```


### Test Stat X GeomSF w/ `make_constructor()`

We can define user facing function with `make_constructor`. However, for greater usability coord_sf(), is included in the user-facing geom_sf() function, and we will want to do this too. 

```{r}
geom_au_state_no_coords <- 
  make_constructor(GeomSf, 
                   stat = StatRegion0, 
                   ref_data = australia_state_ref)

geom_au_state_text_no_coords <- 
  make_constructor(GeomText, 
                   stat = StatRegion0, 
                   ref_data = australia_state_ref)

crs_au_states <- sf::st_crs(sf_oz)

ggplot(au_states) +
  aes(state_name = state) + 
  geom_au_state_no_coords() + 
  geom_au_state_text_no_coords(check_overlap = T) + 
  aes(fill = pop) +
  coord_sf(crs = crs_au_states)
```


```{r}
# Note, in fact, errors w/o coord_sf, and we want to choose the right crs
ggplot(au_states) +
   aes(state_name = state) + 
   geom_au_state_no_coords() 

```



## Step 3.  Create `write_geom_region()` and friends and test

<details Arrow to see details>

```{r old_write_geom_region, eval = T}
# when the crs is NULL, st_crs unfortunately returns, na, but we would like it to return NULL
geom_region0 <- function (mapping = aes(), data = NULL, stat = StatRegion0, position = "identity", 
    na.rm = FALSE, show.legend = NA, inherit.aes = TRUE, ref_data, ...) 
{
    c(layer_sf(geom = GeomSf, data = data, mapping = mapping, 
        stat = stat, position = position, show.legend = show.legend, 
        inherit.aes = inherit.aes, params = rlang::list2(na.rm = na.rm, ref_data = ref_data, 
            ...)), 
    coord_sf(crs = st_crs_mod(ref_data)))
}


#' @export
write_geom_region0 <- function(ref_data, required_aes = NULL){

  modified_fun <- geom_region

  formals(modified_fun)$ref_data <- substitute(ref_data)

  return(modified_fun)

}


# all the arguments should be passed
stamp_region0 <- function (mapping = aes(), data = ref_data, stat = StatRegion0, position = "identity", 
    na.rm = FALSE, show.legend = NA, inherit.aes = FALSE, ref_data, ...) 
{
    c(layer_sf(geom = GeomSf, data = data, mapping = mapping, 
        stat = stat, position = position, show.legend = show.legend, 
        inherit.aes = inherit.aes, params = rlang::list2(na.rm = na.rm, ref_data = ref_data, stamp = T,
            ...)), 
    coord_sf(crs = st_crs_mod(ref_data)))
}



#' @export
write_stamp_region0 <- function(ref_data, required_aes = NULL){

  modified_function <- stamp_region0

formals(modified_function)$ref_data <- substitute(ref_data)

return(modified_function)

}


# all the arguments should be passed
geom_region_text0 <- function (mapping = aes(), data = NULL, stat = StatRegion0,
                              position = "identity", 
    na.rm = FALSE, show.legend = NA, inherit.aes = TRUE, ref_data, ...) 
{
    c(layer_sf(geom = GeomText, data = data, mapping = mapping, 
        stat = stat, position = position, show.legend = show.legend, 
        inherit.aes = inherit.aes, params = rlang::list2(na.rm = na.rm, ref_data = ref_data, 
            ...)), 
    coord_sf(crs = st_crs_mod(ref_data)))
}



#' @export
write_geom_region_text0 <- function(ref_data, required_aes = NULL){

  modified_function <- geom_region_text0

  formals(modified_function)$ref_data <- substitute(ref_data)

return(modified_function)

}



# all the arguments should be passed
# all the arguments should be passed
stamp_region_text0 <- function (mapping = aes(), data = ref_data, stat = StatRegion0,
                              position = "identity", 
    na.rm = FALSE, show.legend = NA, inherit.aes = FALSE, ref_data, ...) 
{
    c(layer_sf(geom = GeomText, data = data, mapping = mapping, 
        stat = stat, position = position, show.legend = show.legend, 
        inherit.aes = inherit.aes, params = rlang::list2(na.rm = na.rm, ref_data = ref_data, stamp = T,
            ...)), 
    coord_sf(crs = st_crs_mod(ref_data)))
}


#' @export
write_stamp_region_text0 <- function(ref_data, required_aes = NULL){

  modified_function <- stamp_region_text0

  formals(modified_function)$ref_data <- substitute(ref_data)

  return(modified_function)

}
```


## Some exploratory work done with `make_constructor`



```{r geom_region0, eval = T, error = T}
GeomSfStamp <- ggplot2::ggproto("GeomSfStamp", ggplot2::GeomSf,
                       default_aes = modifyList(ggplot2::GeomSf$default_aes, 
                                                ggplot2::aes(fill = ggplot2::from_theme(scales::col_mix(ink, paper, 0.8))), 
                                                keep.null = T))

geom_region00 <- ggplot2::make_constructor(ggplot2::GeomSf, stat = StatRegion0) 
stamp_region00 <- ggplot2::make_constructor(GeomSfStamp, stat = StatRegion0, stamp = T, inherit.aes = F)
geom_region_text00 <- ggplot2::make_constructor(ggplot2::GeomText, stat = StatRegion0)
stamp_region_text00 <- ggplot2::make_constructor(ggplot2::GeomText, stat = StatRegion0, stamp = T, inherit.aes = F)
```


## Step 3.b Make geom_region (and friends) that brings along coords_sf() (currently not fully argumented)

```{r}
# all the arguments *should* be passed, but this for the sake of demo
geom_regionX <- function(..., ref_data){
  c(geom_region00(..., ref_data = ref_data), 
    coord_sf(crs = st_crs_mod(ref_data)))
}

# all the arguments *should* be passed, but this for the sake of demo
stamp_regionX <- function(..., ref_data){
  c(stamp_region00(..., ref_data = ref_data), 
    coord_sf(crs = st_crs_mod(ref_data)))
}

geom_region_textX <- function(..., ref_data){
  c(geom_region_text00(..., ref_data = ref_data), 
    coord_sf(crs = st_crs_mod(ref_data)))
}

stamp_region_textX <- function(..., ref_data){
  c(stamp_region_text00(..., ref_data = ref_data), 
    coord_sf(crs = st_crs_mod(ref_data)))
}
```

## Step 4. demo region-specific user-facing functions... just for demonstration purposes!

Not argumented. We'll return to this later.

```{r}
# all arguments above that should be passed, could be passed, or, 
geom_au_states <- function(...){geom_regionX(..., ref_data = australia_state_ref)}
stamp_au_states <- function(...){stamp_regionX(..., ref_data = australia_state_ref)}
geom_au_states_text <- function(...){geom_region_textX(..., ref_data = australia_state_ref)}
stamp_au_states_text <- function(...){stamp_region_textX(..., ref_data = australia_state_ref)}
```

### Test

```{r}
au_states |> 
  sample_n(5) |> 
  ggplot() +
  aes(state_name = state) + 
  stamp_au_states() +
  geom_au_states()

au_states |>
ggplot() +
  aes(state_name = state, fill = pop) + 
  geom_au_states(color = "white") +
  stamp_au_states(keep = "Tasmania", 
                  fill = "cadetblue1"
                  ) +
  stamp_au_states(keep = "Queensland") + 
  geom_au_states_text(color = "red",
                      check_overlap = T)  


last_plot() + 
  geom_au_states(fill = "cadetblue1")

au_states |>
  ggplot() +
  aes(state_name = state) + 
  geom_au_states(aes(fill = pop)) + 
  stamp_au_states(color = "red", keep = "Western Australia",
                  fill = NA, linewidth = 3)
```


</details>


