---
output: 
  github_document:
    toc: TRUE
    toc_depth: 2
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>", 
  message = F,
  warning = F
)
```

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

# {ggregions}

Note: foundation work for this package has been conducted in many experiments 
([1](https://evamaerey.github.io/mytidytuesday/2025-09-15-make_constructor_sf/make_constructor_sf.html),
[2](https://evamaerey.github.io/mytidytuesday/2024-07-12-sf-experiment-fastStat/sf-experiment-fastStat.html),
[3](https://evamaerey.github.io/mytidytuesday/2024-06-13-usmaps-ggusa/usmaps-ggusa.html ),
[4](https://evamaerey.github.io/mytidytuesday/2024-06-13-usmaps-ggusa/usmaps-ggusa-Stat-and-geom-sf.html ),
[5](https://evamaerey.github.io/mytidytuesday/2024-04-08-statsf_errors/statsf_errors.html ),
[6](https://evamaerey.github.io/mytidytuesday/2023-03-12-ggbrain-seg-sf/ggbrain_seg_sf.html ),
[7](https://evamaerey.github.io/mytidytuesday/2023-03-12-ggbrain-seg-sf/ggbrain_seg_dk_sf.html ),
[8](https://evamaerey.github.io/mytidytuesday/2023-03-10-ggfips/ggfips_w_sf.html ),
[9](https://evamaerey.github.io/mytidytuesday/2023-03-10-ggfips/ggfips.html ),
[10](https://evamaerey.github.io/mytidytuesday/2023-03-10-brain2/brain2.html ),
[11](https://evamaerey.github.io/mytidytuesday/2023-03-09-nc-fips/nc-fips.html ),
[12](https://evamaerey.github.io/mytidytuesday/2023-03-07-extending-your-ability/extending_your_ability.html ),
[13](https://evamaerey.github.io/mytidytuesday/2023-03-06-us-states/us_states.html ) ),
 [a number of discussions](https://github.com/search?q=repo%3Aggplot2-extenders%2Fggplot-extension-club+sf2stat&type=discussions) and in conversation based on my [talk](https://evamaerey.github.io/mytidytuesday/2023-04-13-ggcircle-pack-talk/circle_pack_and_beyond_talk.html#86) in the ggplot2 extenders group,
and a couple of experimental packages ([sf2stat](https://github.com/EvaMaeRey/sf2stat) and [ggsomewhere](https://github.com/EvaMaeRey/ggsomewhere)).



# Motivations

> I am told there are people who do not care for maps, and I find it hard to believe. - Robert Louis Stevenson

> 'My teacher, Mr. Jayson, says a map is a picture of someplace from above.  It's like flying over that spot in an airplane.' - 'Lisa' in Loreen Leedy's in 'Mappying Penny's World'

> I hate when I work with spatial data that so much of my mindshare goes to thinking about 'how do I wrangle this this map and what on earth is going on' versus the actual meaning behind what I'm trying to map. - Emily Riederer

Maps are some of the most compelling and intuitive data visualizations, and they are some of the first visualizations people will come across in their lives.  There is great support for mapping in R and in the data vizualization library {ggplot2} which touches many R packages.  

However, producing a simple map, like a choropleth, can feel much harder than producing other plots, like a scatterplot for example in ggplot2.  A scatter plot is produced by:

```r 
ggplot(scatter_data) + # data
   aes(x = x, y = y) + # position and other visual channel mapping
   geom_point()        # layer 
```

In words, you might say, 'used this data, and points, and the positions are defined by x and y'

This proposal posits that by investing in methods to easily create new geo-specific ggplot2 layer extensions (geom_*s), producing maps with ggplot2 can become as intuitive as creating other plot as well as reading the maps themselves. About specifying a map, you might say, 'use this data, draw regions, position and color fill according to var1 and var2.'

The proposed regions mapping might look something like this:

```r
ggplot2(mapping_data) +             # data
   aes(region_id = var1_region_id,  # position mapping 
       fill = var2) +               # and other visual channel mapping
   geom_region()                    # layer
```

> Your API feels like it actually puts the "question" at the center.. - Emily Riederer


# What {ggregions} will deliver

## Effortlessly write new geom_* region-specific functions

The {ggregions} package (in proof-of-concept phase) will allow geographic data package writers to quickly create easy-to-use ggplot2 layer extensions that adhere to the ggplot2 'grammar'.

First, the authors will prepare a reference dataset that contains a [geometry](https://ggplot2.tidyverse.org/reference/ggsf.html#geometry-aesthetic), sfc-containing column as well as columns of allowed identifiers for the regions of interest.  For example, the identifiers from in `us_states_ref` from `usmapdata::us_map()`, are `state_name`, `fips` and `state_abbr`. 


```{r, eval = T}
library(dplyr)

# Step 1. prep some of their geo reference data
us_states_ref <- usmapdata::us_map() |>
  select( state_name = full, # state_name will positional aes like x or y 
          state_abbr = abbr, 
          fips,
          geometry = geom,  # one column with geometry is required
         )  # state_abbr will also be optional positional aes like x or y

head(us_states_ref)
```


Then, authors will be able use `write_geom_region_locale()` to create a regions-specific ggplot2 layer, for example `geom_us_state()` below, which can be made available in their geo packages.

```{r}
# Step 2. use `write_geom_region_locale`
library(ggregions)
geom_us_state <- write_geom_region_locale(ref_data = us_states_ref)
```

## Deliver intuitive, newcomer-welcoming spatial viz experience

By including the newly specified `geom_` function in their geo data package (perhaps instead or in addition to a convenience wrapper that are typical of geo packages) the package *users* will be able to use the` data + aes + geom` formulation for defining their plots.  Instead of positional aesthetics like x, y or geometry, a region identifier like `state_name` or `fips` can be used.  Internally, the regions `geom_*` will translate between place name or other identifier to the necessary borders (`sfc`s).

Below is some tabular data with geographic region identifiers, but no boundaries:

```{r, eval = T}
us_income <- tidyr::us_rent_income |> 
  filter(variable == "income") |>
  rename(income = estimate)

head(us_income)
```

```{r, echo = F}
library(ggplot2)
(theme_gray() + 
  theme(palette.colour.continuous = "magma")) |> 
  set_theme()
```

Even though there is no boundary information, analysts can map this data with the familiar  data + aesthetic mapping + layer syntax that makes ggplot2 easy to read, write, and reason about.

```{r}
library(ggplot2)
library(usmapdata)

ggplot(data = us_income) + 
  aes(state_name = NAME, 
      fill = income) + 
  geom_us_state()
```

Packages that might take avantage of ggregion's functionality include
[afrimapr](https://afrimapr.github.io/afrimapr.website/), [geobr](https://ipeagit.github.io/geobr/), [cancensus](https://mountainmath.github.io/cancensus/index.html), [chilemapas](https://pacha.dev/chilemapas/), [RCzechia](https://github.com/jlacko/RCzechia), [geofi](https://ropengov.github.io/geofi/)
, [geokz](https://github.com/arodionoff/geokz), [mapsPERU](https://github.com/musajajorge/mapsPERU) and [geoidep](https://geografo.pe/geoidep/index.html), [mapSpain](https://github.com/rOpenSpain/mapSpain/), [geographr](https://github.com/britishredcrosssociety/geographr), [geouy](https://github.com/RichDeto/geouy), [tigris](https://github.com/walkerke/tigris),
and [rgeoboundaries](https://github.com/wmgeolab/rgeoboundaries).
 Since these geo packages have many users, the impacts of `{ggregions}` could be significant.


# Further disucssion and implementation demonstrations, ideas


## Status Quo: Inconsistent APIs or leave the data manipulation to the user.

```{r, message=F, warning=F, fig.show='hide'}
library(tidyverse)
library(ozmaps)

# some data that we want to viz
au_states <- tribble(~state, ~pop,
        "Victoria", 1,
        "Queensland", 2,
        "New South Wales", 3,
        "Western Australia", 4,
        "Northern Territory", 5,
        "Tasmania", 6,
        "South Australia", 7,
        "Australian Capital Territory", 8,
        "Other Territories", 9
        )

sf_oz <- ozmap("states")

sf_oz |> names()

sf_oz |>
  full_join(au_states, by = join_by(NAME == state )) |> 
  ggplot() + 
  aes(geometry = geometry, fill = pop) + 
  geom_sf()
```



# Alternative: Toward region-specific layer-writing 


## Step 1. Define compute

We'll look at the Australian states case as we go, but the ref_data argument will keep thing generic.

```{r}
australia_state_ref <- sf_oz |>
  select(state_name = NAME)
```


```{r}
compute_panel_regions <- function(data, scales, ref_data, keep = NULL, drop = NULL, stamp = F){

ref_data$id <- ref_data[1][[1]]

if(!is.null(keep)){ref_data <- ref_data |> dplyr::filter(id %in% keep)}
if(!is.null(drop)){ref_data <- ref_data |> dplyr::filter(!(id %in% drop))}

ref_data <- ref_data |> 
    ggplot2::StatSf$compute_panel(coord = ggplot2::CoordSf) |> # add bounding boxes xmin xmax etc
    ggplot2::StatSfCoordinates$compute_group(coord = ggplot2::CoordSf) # add x and y centers

if(!stamp){ ref_data |> dplyr::inner_join(data) } else { ref_data }

}
```


### Test Compute

```{r}
au_states |> 
  rename(state_name = state) |>
  compute_panel_regions(ref_data = australia_state_ref)
```

To do/consider

-[]warnings about not plot data that fails to find a match in ref data
-[]allowing multiple geo reference data sets with different resolutions?
-[]more flexibility on specifying the id column (now it is the first column in the dataset)
-[]other?

## Step 2. Define Stat

```{r StatRegion}
# compute just reproduced verbatim from above for packaging purposes
compute_panel_regions <- function(data, scales, ref_data, keep = NULL, 
                                  drop = NULL, stamp = F){

  ref_data$id <- ref_data[1][[1]]

  if(!is.null(keep)){ref_data <- ref_data |> dplyr::filter(id %in% keep)}
  if(!is.null(drop)){ref_data <- ref_data |> dplyr::filter(!(id %in% drop))}

  ref_data <- ref_data |> 
    ggplot2::StatSf$compute_panel(coord = ggplot2::CoordSf) |>
    ggplot2::StatSfCoordinates$compute_group(coord = ggplot2::CoordSf)

  if(!stamp){ ref_data |> dplyr::inner_join(data) } else { ref_data }

}


StatRegion <- ggplot2::ggproto("StatRegion",
                      ggplot2::Stat,
                      compute_panel = compute_panel_regions,
                      default_aes = ggplot2::aes(label = ggplot2::after_stat(id)))
```

### To do/consider


-[]Stat won't warn about required_aes, since none are declared.  We anticipate allowed aes to be quite varied.  How otherwise can this be addressed?
-[]Think through, experiment if compute_group might also work, and if it is preferable

### Test Stat

```{r}
ggplot(au_states) +
  aes(state_name = state) +
  geom_sf(stat = StatRegion, ref_data = australia_state_ref) + 
  geom_text(stat = StatRegion, ref_data = australia_state_ref) +
  aes(fill = pop)
```


### Test Stat X GeomSF w/ `make_constructor()`

We can define user facing function with `make_constructor`. However, for greater usability coord_sf(), is included in the user-facing geom_sf() function, and we will want to do this too. 

```{r}
geom_au_state_no_coords <- 
  make_constructor(GeomSf, 
                   stat = StatRegion, 
                   ref_data = australia_state_ref)

geom_au_state_text_no_coords <- 
  make_constructor(GeomText, 
                   stat = StatRegion, 
                   ref_data = australia_state_ref)

crs_au_states <- sf::st_crs(sf_oz)

ggplot(au_states) +
  aes(state_name = state) + 
  geom_au_state_no_coords() + 
  geom_au_state_text_no_coords(check_overlap = T, size = 2) + 
  aes(fill = pop) +
  coord_sf(crs = crs_au_states)


# Note, in fact, errors w/o coord_sf, and we want to choose the right crs
# ggplot(au_states) +
#   aes(state_name = state) + 
#   geom_au_state_no_coords() 

```


## Step 3.a Make geom_region0 (no coords) and friends with `make_constructor`


```{r geom_region0, eval = T, error = T}
geom_region0 <- ggplot2::make_constructor(ggplot2::GeomSf, stat = StatRegion) 
stamp_region0 <- ggplot2::make_constructor(ggplot2::GeomSf, stat = StatRegion, stamp = T, inherit.aes = F)
geom_region_text0 <- ggplot2::make_constructor(ggplot2::GeomText, stat = StatRegion)
stamp_region_text0 <- ggplot2::make_constructor(ggplot2::GeomText, stat = StatRegion, stamp = T, inherit.aes = F)
```


## Step 3.b Make geom_region (and friends) that brings along coords_sf() (currently not fully argumented)

```{r}
# all the arguments *should* be passed, but this for the sake of demo
geom_region <- function(..., ref_data){
  c(geom_region0(..., ref_data = ref_data), 
    coord_sf(crs = sf::st_crs(ref_data)))
}

# all the arguments *should* be passed, but this for the sake of demo
stamp_region <- function(..., ref_data){
  c(stamp_region0(..., ref_data = ref_data), 
    coord_sf(crs = sf::st_crs(ref_data)))
}

geom_region_text <- function(..., ref_data){
  c(geom_region_text0(..., ref_data = ref_data), 
    coord_sf(crs = sf::st_crs(ref_data)))
}

stamp_region_text <- function(..., ref_data){
  c(stamp_region_text0(..., ref_data = ref_data), 
    coord_sf(crs = sf::st_crs(ref_data)))
}
```

## Step 4. make region-specific user-facing functions!

Not argumented. We'll return to this later.

```{r}
# all arguments above that should be passed, could be passed, or, 
geom_au_states <- function(...){geom_region(..., ref_data = australia_state_ref)}
stamp_au_states <- function(...){geom_region(..., ref_data = australia_state_ref)}
geom_au_states_text <- function(...){geom_region_text(..., ref_data = australia_state_ref)}
stamp_au_states_text <- function(...){geom_region_text(..., ref_data = australia_state_ref)}
```

### Test

```{r}
au_states |>
ggplot() +
  aes(state_name = state,
      fill = pop) + 
  geom_au_states() +
  geom_au_states(stamp = T,
              keep = "Tasmania", 
              fill = "orange") +
  geom_au_states_text()  

au_states |>
  ggplot() +
  aes(state_name = state) + 
  geom_au_states(aes(fill = pop)) + 
  stamp_au_states(color = "red", keep = "Western Australia")
```




## nc test

<detail>

```{r}
nc_ref <- sf::st_read(system.file("shape/nc.shp", package="sf")) |>
  select(county_name = NAME, fips = FIPS)

geom_nc_county <- function(...){geom_region(..., ref_data = nc_ref)}

nc_data <- sf::st_read(system.file("shape/nc.shp", package="sf")) |>
  sf::st_drop_geometry()

nc_data |>
  ggplot() +
  aes(county_name = NAME,
      fill = AREA) + 
  geom_nc_county()
```

</detail>



## To do... fully argumenting function with the approach used below (or more compact) for `write_geom_region_locale()`

```{r write_geom_region_locale, eval = T}
# all the arguments should be passed
geom_region <- function(mapping = NULL, data = NULL, stat = StatRegion, position = "identity", 
                        ..., legend = NULL, lineend = "butt", linejoin = "round", 
    linemitre = 10, arrow = NULL, arrow.fill = NULL, na.rm = FALSE, 
    show.legend = NA, inherit.aes = TRUE, ref_data){
  
  c(geom_region0(mapping = mapping, data = data, stat = stat, 
        position = position, show.legend = show.legend, inherit.aes = inherit.aes, ref_data = ref_data,
        params = rlang::list2(na.rm = na.rm, legend = legend, lineend = lineend, 
            linejoin = linejoin, linemitre = linemitre, arrow = arrow, 
            arrow.fill = arrow.fill, ...)), 
    coord_sf(crs = sf::st_crs(ref_data)))
  
}

#' @export
write_geom_region_locale <- function(ref_data = australia_state_ref){

  geom_region_local <- geom_region

formals(geom_region_local)$ref_data <- substitute(ref_data)

return(geom_region_local)

}
```


## Australia example w/ `write_geom_region_locale()`

```{r}
geom_au_state <- write_geom_region_locale(ref_data = australia_state_ref)

geom_au_state

au_states |> 
  ggplot() + 
  aes(state_name = state, fill = pop) + 
  geom_au_state()

geom_au_state
```

## US state example w/ `write_geom_region_locale()`

```{r}
us_states_ref <- usmapdata::us_map() |>
  rename(geometry = geom,
         state_name = full,
         state_abbr = abbr) |>
  select(state_name, everything())
  
geom_us_state <- write_geom_region_locale(ref_data = us_states_ref)

us_rent_income |> 
  filter(variable == "income") |>
  ggplot() + 
  aes(state_name = NAME, fill = estimate) + 
  geom_us_state() + 
  scale_fill_viridis_c(option = "magma")
```


# Minimal Packaging


### Do once

```{r, eval = F}
devtools::create(".")
usethis::use_mit_license()
usethis::use_lifecycle_badge("experimental")
```

### In manually and in readme

```{r}
knitrExtra::chunk_to_dir("StatRegion")
knitrExtra::chunk_to_dir("geom_region0")
knitrExtra::chunk_to_dir("write_geom_region_locale")

usethis::use_package("ggplot2")
usethis::use_package("sf")
usethis::use_package("dplyr")

devtools::document()
```


### Do manually (time consuming)

```{r, eval = F}
devtools::check(".")
devtools::install(pkg = ".", upgrade = "never")
```



